---
title: "Titanic"
author: "Michał Makowski"
output: 
  html_notebook: 
    toc: yes
    toc_depth: 2
    highlight: pygments
    theme: cosmo
  html_document: 
    toc: yes
    toc_depth: 2
    highlight: pygments
    theme: cosmo
    df_print: kable
---

# Introduction

This is Markdown notebook which documents my work over *Titanic* data from *Kaggle*. It is similiar to a Jupyter notebook which I've done in python.


# Setup and preparation

Data are provided in two datasets - **test** and **train**. First provides labelled data, second is used for *Kaggle* submission and testing. We loaded everything into memory and briefly checked the structure of the data. 

```{r setup, message = FALSE, collapse = TRUE, strip.white = TRUE}

knitr::opts_chunk$set(results = "hold", message = FALSE, strip.white = TRUE, warning = FALSE)

loadLibraries <- c("tidyverse", "cowplot", "lattice", "GGally", "corrplot", "pipeR", "caTools",
				   "themis", "ranger", "randomForest", "glmnet", "vip",
				   "rsample", "tidymodels", "parsnip", "yardstick", "tune", "dials")

installedLibraries <- loadLibraries[!loadLibraries %in% installed.packages()]

for(lib in installedLibraries) 
	install.packages(lib, dependences = TRUE)
sapply(loadLibraries, require, character = TRUE)

# for replication purposes
# initialSeed	<- as.integer(Sys.time())
# theSeed		<- initialSeed %% 100000
# print(theSeed) # 41428

set.seed(41428)
theme_set(theme_bw()) # ggplot theme

```


```{r dataImport, warning=TRUE}

test  <- read_csv("01 Data/train")
train <- read_csv("01 Data/test.csv")

# head(customers)
# head(pilot)
# head(decisions)

# identical(customers, customers[complete.cases(customers),])
# all(customers == customers[complete.cases(customers),]) # :)
# all(pilot == pilot[complete.cases(pilot),])
# both TRUE

```