---
title: "Titanic"
author: "Michał Makowski"
output: 
  html_notebook: 
    toc: yes
    toc_depth: 2
    highlight: pygments
    theme: cosmo
  html_document: 
    toc: yes
    toc_depth: 2
    highlight: pygments
    theme: cosmo
    df_print: kable
---

# Introduction

This is Markdown notebook which documents my work over *Titanic* data from *Kaggle*. It is similiar to a Jupyter notebook which I've done in python.


# Setup and preparation

Data are provided in two datasets - **test** and **train**. First provides labelled data, second is used for *Kaggle* submission and testing. We loaded everything into memory and briefly checked the structure of the data. 

```{r setup, message = FALSE, collapse = TRUE, strip.white = TRUE}

knitr::opts_chunk$set(results = "hold", message = FALSE, strip.white = TRUE, warning = FALSE)

loadLibraries <- c("tidyverse", "cowplot", "lattice", "GGally", "corrplot", "pipeR", "caTools",
				   "themis", "ranger", "randomForest", "glmnet", "vip",
				   "rsample", "tidymodels", "parsnip", "yardstick", "tune", "dials")

installedLibraries <- loadLibraries[!loadLibraries %in% installed.packages()]

for(lib in installedLibraries) 
	install.packages(lib, dependences = TRUE)
sapply(loadLibraries, require, character = TRUE)

# for replication purposes
# initialSeed	<- as.integer(Sys.time())
# theSeed		<- initialSeed %% 100000
# print(theSeed) # 41428

set.seed(41428)
theme_set(theme_bw()) # ggplot theme

```


```{r dataImport, warning=TRUE}

train <- read_csv("01 Data/test.csv")
test  <- read_csv("01 Data/train.csv")

head(train)
head(test)

identical(train, train[complete.cases(train),])
identical(test, test[complete.cases(test),])

```
```{r trainPrep}

train$Pclass	<- train$Pclass %>% as.character() %>% parse_factor(sort(unique(.))) # alphabetical order of factors
train$Sex		<- train$Sex %>% parse_factor(sort(unique(.))) # ditto
train$Age		<- as.integer(train$Age)
train$SibSp		<- as.integer(train$SibSp)
train$Parch	    <- as.integer(train$Parch)
train$Embarked	<- train$Embarked %>% parse_factor(sort(unique(.))) # ditto
# str(train)
# summary(train)

```

```{r testPrep}

test$Pclass		<- test$Pclass %>% as.character() %>% parse_factor(sort(unique(.))) # alphabetical order of factors
test$Sex		<- test$Sex %>% parse_factor(sort(unique(.))) # ditto
test$Age		<- as.integer(test$Age)
test$SibSp		<- as.integer(test$SibSp)
test$Parch	    <- as.integer(test$Parch)
test$Embarked	<- test$Embarked %>% parse_factor(sort(unique(.))) # ditto
# str(test)
# summary(test)

```


# Data quality

## Customer dataset and Pilot dataset structure similarity

As we wanted to build the model on the **treain** dataset and use the **test** dataset for predictions, we checked if there are any significant differences in distributions of variables between those sets. 

*We could had trusted our data provider, but at least simple visual inspection should be done.*

### Plots

We do not present every plot, as distributions in both groups were pairwise similar.

```{r continousSimilarityEverything}

# continuous variables
everyCase <- bind_rows(mutate(train, 
							  test = FALSE),
					   mutate(select(test, -Survived),
					   	   test = TRUE))
everyCase %>%
	select(where(is.double)) %>%
	colnames() %>%
	sapply(
		function(X)
		{
			ggplot(everyCase) +
				geom_area(aes(x = get(X), 
							  y = ..density.., 
							  fill = test,
							  colour = test),
						  alpha = 0.25,
						  stat = "bin",
						  bins = 93, # Rice rule
						  position = "identity") +
				labs(x = X) +
				theme(legend.position = c(0.8, 0.8)) -> plot1
				
			ggplot(everyCase) +
				geom_violin(aes(x = test,
								y = get(X),
								fill = test),
							alpha = 0.15) +
				geom_boxplot(aes(x = test, 
								 y = get(X),
								 fill = test),
							 alpha = 0.75) +
				labs(y = X) +
				guides(fill = "none") -> plot2
			
			
			plot_grid(plot1, plot2,  rel_widths = c(5, 3))
		},
		simplify = FALSE,
		USE.NAMES = TRUE
	) -> plotListContinous

plotListContinous$Fare

```

```{r discreteSimilarityEverything}

# discrete variables
everyCase %>%
	select(where(is.integer)) %>%
	colnames() %>%
	sapply(
		function(X)
		{
			ggplot(everyCase) +
				geom_bar(aes(x = get(X), 
							 y = ..prop..,
							 fill = test,
							 colour = test),
							 alpha = 0.4,
							 position = "identity") +
				labs(x = X) +
				theme(legend.position = c(0.9, 0.8)) -> plot1
				
			ggplot(everyCase) +
				geom_boxplot(aes(x = test, 
								 y = get(X),
								 fill = test),
							 alpha = 0.75) +
				labs(y = X) +
				guides(fill = "none") -> plot2
			
			plot_grid(plot1, plot2,  rel_widths = c(5, 3))
		},
		simplify = FALSE,
		USE.NAMES = TRUE
	) -> plotListDiscrete

everyCase %>%
	select(where(is.factor)) %>%
	colnames() %>%
	sapply(
		function(X)
		{
			ggplot(everyCase) +
				geom_bar(aes(x = get(X),
							 y =  ..prop..,
							 group = test,
							 fill = test,
							 colour = test),
						 alpha = 0.25,
						 stat = "count",
						 position = "identity") +
				labs(x = X, 
					 y = "Probability") +
				theme(legend.position = c(0.8, 0.8))
		},
		simplify = FALSE,
		USE.NAMES = TRUE
	) -> plotListFactor

plotListDiscrete$Age
plotListFactor$Pclass

# table(everyCase$test, everyCase$Sex) %>%
# 	proportions(1)
# Proportions do not differ to much
# table(everyCase$test, everyCase$SibSp) %>%
# 	proportions(1)
# Ditto, results do not differ much.

```

The brief introductory analysis showed that both sets (**train** and **test**) are almost similar in terms of variables' distributions. To confirm that there could be done more analysis which might reveal some differences or conditional dependencies. As the task should not dive to deeply into the problem, we assumed that variables are not highly correlated (what was somehow proven by plots described in next paragraph).

Correlation analysis is quite useful when we want to interpret results and add some **domain** knowledge to the problem. Scatter plots are great tools for that, they can be produced by **pairs** or **ggpairs** functions.


```{r ggPairsEverything}

filter(everyCase, test == FALSE) %>%
	select(-Name, -Ticket, -Cabin, -test) %>%
	ggpairs()
filter(everyCase, test == TRUE) %>%
	select(-Name, -Ticket, -Cabin, -test) %>%
	ggpairs()

```